---
title: "Mini-Project #01: Gourmet Cheeseburgers Across the Globe: Exploring the Most Popular Programming on Netflix"
author: "Haley Vargas"
date: today
format:
  html:
    code-fold: true
    code-summary: "Show the code"
    toc: true
    toc-depth: 3
    toc-title: "📚 Contents"
    smooth-scroll: true
execute:
  echo: true
  warning: false
  message: false
---
# Acquiring the Data
```{r}
if(!dir.exists(file.path("data", "mp01"))){
    dir.create(file.path("data", "mp01"), showWarnings=FALSE, recursive=TRUE)
}

GLOBAL_TOP_10_FILENAME <- file.path("data", "mp01", "global_top10_alltime.csv")

if(!file.exists(GLOBAL_TOP_10_FILENAME)){
    download.file("https://www.netflix.com/tudum/top10/data/all-weeks-global.tsv", 
                  destfile=GLOBAL_TOP_10_FILENAME)
}

COUNTRY_TOP_10_FILENAME <- file.path("data", "mp01", "country_top10_alltime.csv")

if(!file.exists(COUNTRY_TOP_10_FILENAME)){
    download.file("https://www.netflix.com/tudum/top10/data/all-weeks-countries.tsv", 
                  destfile=COUNTRY_TOP_10_FILENAME)
}
```

# Data Import and Preparation
```{r}
if(!require("tidyverse")) install.packages("tidyverse")
library(readr)
library(dplyr)
GLOBAL_TOP_10 <- read_tsv(GLOBAL_TOP_10_FILENAME)
str(GLOBAL_TOP_10)
glimpse(GLOBAL_TOP_10)
```

## Task 2: Data Cleaning
```{r}
GLOBAL_TOP_10 <- GLOBAL_TOP_10 |>
  mutate(season_title = if_else(season_title == "N/A", NA_character_, season_title))
```

## Task 3: Data Import
```{r}
COUNTRY_TOP_10 <- read_tsv(COUNTRY_TOP_10_FILENAME, na = "N/A")
glimpse(COUNTRY_TOP_10)
view(COUNTRY_TOP_10)
```

# Initial Data Exploration
```{r}
library(DT)
GLOBAL_TOP_10 |>
  head(n = 20) |>
  datatable(options = list(searching = FALSE, info = FALSE))
```
```{r}
library(stringr)
format_titles <- function(df) {
  colnames(df) <- str_replace_all(colnames(df), "_", " ") |> str_to_title()
  df
}

GLOBAL_TOP_10 |>
  format_titles() |>
  head(n = 20) |>
  datatable(options = list(searching = FALSE, info = FALSE)) |>
  formatRound(c('Weekly Hours Viewed', 'Weekly Views'))
```
```{r}
GLOBAL_TOP_10 |>
  select(-season_title) |>
  format_titles() |>
  head(n = 20) |>
  datatable(options = list(searching = FALSE, info = FALSE)) |>
  formatRound(c('Weekly Hours Viewed', 'Weekly Views'))
```
```{r}
GLOBAL_TOP_10 |>
  mutate(`runtime_(minutes)` = round(60 * runtime)) |>
  select(-season_title, -runtime) |>
  format_titles() |>
  head(n = 20) |>
  datatable(options = list(searching = FALSE, info = FALSE)) |>
  formatRound(c('Weekly Hours Viewed', 'Weekly Views'))
```
## Task 4: Exploratory Questions

### Question 1: How many different countries does Netflix operate in? (You can use the viewing history as a proxy for countries in which Netflix operates.)

```{r}
COUNTRY_TOP_10 |>
  summarize("Number of different countries Netflix operates in" = n_distinct(country_name))
```

### Question 2: Which non-English-language film has spent the most cumulative weeks in the global top 10? How many weeks did it spend?

```{r}
top_film <- GLOBAL_TOP_10 |> filter(category == "Films (Non-English)") |> arrange(desc(cumulative_weeks_in_top_10)) |>  slice(1)
print(top_film)
top_film$cumulative_weeks_in_top_10
top_film$show_title
```

### Question 3: What is the longest film (English or non-English) to have ever appeared in the Netflix global Top 10? How long is it in minutes?
```{r}
longest_film <- GLOBAL_TOP_10 |> filter(category %in% c("Films (English)", "Films (Non-English)")) |> arrange(desc(runtime)) |> slice(1)
longest_film$show_title
longest_film$runtime
```

### Question 4: For each of the four categories, what program has the most total hours of global viewership?
```{r}
top_programs <- GLOBAL_TOP_10 |>
  group_by(category, show_title) |>                     # group by category and title
  summarise(total_hours = sum(weekly_hours_viewed, na.rm = TRUE)) |>  # sum hours viewed
  slice_max(order_by = total_hours, n = 1) %>%           # take top program per category
  ungroup()
print(top_programs)
```

### Question 5: Which TV show had the longest run in a country’s Top 10? How long was this run and in what country did it occur?
```{r}
longest_run <- COUNTRY_TOP_10 |> filter(category == "TV") |> arrange(desc(cumulative_weeks_in_top_10)) |>  slice(1)
longest_run %>%
  select(show_title, country_name, cumulative_weeks_in_top_10)
```

### Question 7:  What is the total viewership of the TV show Squid Game? Note that there are three seasons total and we are looking for the total number of hours watched across all seasons.

```{r}
squid_game_total <- GLOBAL_TOP_10 %>%
  filter(show_title == "Squid Game") %>%
  summarize(total_hours = sum(weekly_hours_viewed, na.rm = TRUE))
print(squid_game_total)
```

### Question 8: The movie Red Notice has a runtime of 1 hour and 58 minutes. Approximately how many views did it receive in 2021? Note that Netflix does not provide the weekly_views values that far back in the past, but you can compute it yourself using the total view time and the runtime.
```{r}
library(lubridate)
library(scales)

# runtime in hours
runtime_hours <- 118 / 60  # 118 minutes = 1.9666667 hours

red_notice_hours_2021 <- GLOBAL_TOP_10 %>%
  filter(show_title == "Red Notice", year(week) == 2021) %>%
  summarize(total_hours = sum(weekly_hours_viewed, na.rm = TRUE)) %>%
  pull(total_hours)

approx_views_2021 <- red_notice_hours_2021 / runtime_hours
print(approx_views_2021)
```


