---
title: "Mini-Project #01: Gourmet Cheeseburgers Across the Globe: Exploring the Most Popular Programming on Netflix"
author: "Haley Vargas"
date: today
format:
  html:
    code-fold: true
    code-summary: "Show the code"
    toc: true
    toc-depth: 3
    toc-title: "📚 Contents"
    smooth-scroll: true
execute:
  echo: true
  warning: false
  message: false
---
# Acquiring the Data
Firstly, we will need to acquire the data from [Netflix's Tudum Top 10](https://www.netflix.com/tudum/top10). Specifically, we will be analyzing two separate data files:

* Global Top 10
* Country-wide Top 10
```{r}
if(!dir.exists(file.path("data", "mp01"))){
    dir.create(file.path("data", "mp01"), showWarnings=FALSE, recursive=TRUE)
}

GLOBAL_TOP_10_FILENAME <- file.path("data", "mp01", "global_top10_alltime.csv")

if(!file.exists(GLOBAL_TOP_10_FILENAME)){
    download.file("https://www.netflix.com/tudum/top10/data/all-weeks-global.tsv", 
                  destfile=GLOBAL_TOP_10_FILENAME)
}

COUNTRY_TOP_10_FILENAME <- file.path("data", "mp01", "country_top10_alltime.csv")

if(!file.exists(COUNTRY_TOP_10_FILENAME)){
    download.file("https://www.netflix.com/tudum/top10/data/all-weeks-countries.tsv", 
                  destfile=COUNTRY_TOP_10_FILENAME)
}
```

# Data Import and Preparation
```{r}
if(!require("tidyverse")) install.packages("tidyverse")
library(readr)
library(dplyr)
GLOBAL_TOP_10 <- read_tsv(GLOBAL_TOP_10_FILENAME)
```
After utilizing the read_tsv function, we can use the str() and glimpse() functions to examine its structure:
```{r}
str(GLOBAL_TOP_10)
glimpse(GLOBAL_TOP_10)
```
From these outputs we can see that the GLOBAL_TOP_10 dataframe consists of 9 columns and 8840 rows. It also describes the names of each column.

## Task 2: Data Cleaning
If we notice carefully, we can see that when "season_title" is missing, R is reading these values as strings. To fix this, we will need to use the mutate command. 
```{r}
GLOBAL_TOP_10 <- GLOBAL_TOP_10 |>
  mutate(season_title = if_else(season_title == "N/A", NA_character_, season_title))
```

## Task 3: Data Import
Now we will do the same for COUNTRY_TOP_10:
```{r}
COUNTRY_TOP_10 <- read_tsv(COUNTRY_TOP_10_FILENAME, na = "N/A")
glimpse(COUNTRY_TOP_10)
view(COUNTRY_TOP_10)
```

# Initial Data Exploration
Now lets use the DT package to better visualize the data: 
```{r}
library(DT)
GLOBAL_TOP_10 |>
  head(n = 20) |>
  datatable(options = list(searching = FALSE, info = FALSE))
```
```{r}
library(stringr)
format_titles <- function(df) {
  colnames(df) <- str_replace_all(colnames(df), "_", " ") |> str_to_title()
  df
}

GLOBAL_TOP_10 |>
  format_titles() |>
  head(n = 20) |>
  datatable(options = list(searching = FALSE, info = FALSE)) |>
  formatRound(c('Weekly Hours Viewed', 'Weekly Views'))
```
```{r}
GLOBAL_TOP_10 |>
  select(-season_title) |>
  format_titles() |>
  head(n = 20) |>
  datatable(options = list(searching = FALSE, info = FALSE)) |>
  formatRound(c('Weekly Hours Viewed', 'Weekly Views'))
```
```{r}
GLOBAL_TOP_10 |>
  mutate(`runtime_(minutes)` = round(60 * runtime)) |>
  select(-season_title, -runtime) |>
  format_titles() |>
  head(n = 20) |>
  datatable(options = list(searching = FALSE, info = FALSE)) |>
  formatRound(c('Weekly Hours Viewed', 'Weekly Views'))
```
## Task 4: Exploratory Questions

### Question 1: How many different countries does Netflix operate in? (You can use the viewing history as a proxy for countries in which Netflix operates.)

#### Answer: Netflix operates in **94 countries**

```{r}
COUNTRY_TOP_10 |>
  summarize("Number of different countries Netflix operates in" = n_distinct(country_name))
```

### Question 2: Which non-English-language film has spent the most cumulative weeks in the global top 10? How many weeks did it spend?

#### Answer: *All Quiet on the Western Front* spent the most cumulative weeks in the global top 10. It spent **23** cumulative weeks.

```{r}
top_film <- GLOBAL_TOP_10 |> filter(category == "Films (Non-English)") |> arrange(desc(cumulative_weeks_in_top_10)) |>  slice(1)
print(top_film)
top_film$cumulative_weeks_in_top_10
top_film$show_title
```

### Question 3: What is the longest film (English or non-English) to have ever appeared in the Netflix global Top 10? How long is it in minutes?

#### Answer: *Pushpa 2: The Rule (Reloaded Version* is the longest film to ever appear in the Netflix global top 10, it is approximately 224 minutes long.
```{r}
longest_film <- GLOBAL_TOP_10 |> filter(category %in% c("Films (English)", "Films (Non-English)")) |> arrange(desc(runtime)) |> slice(1)
longest_film$show_title
longest_film_minutes <- longest_film$runtime * 60
round(longest_film_minutes)
```

### Question 4: For each of the four categories, what program has the most total hours of global viewership?
```{r}
top_programs <- GLOBAL_TOP_10 |>
  group_by(category, show_title) |>                     # group by category and title
  summarise(total_hours = sum(weekly_hours_viewed, na.rm = TRUE)) |>  # sum hours viewed
  slice_max(order_by = total_hours, n = 1) %>%           # take top program per category
  ungroup()

datatable(
  top_programs,
  colnames = c("Category", "Show Title", "Total Hours Viewed"),
  options = list(pageLength = 5, autoWidth = TRUE)
) |>
  formatCurrency("total_hours", currency = "", interval = 3, mark = ",")
```

### Question 5: Which TV show had the longest run in a country’s Top 10? How long was this run and in what country did it occur?

#### Answer: *Money Heist* spent the longest run of 127 cumulative weeks in Pakistan. 
```{r}
longest_run <- COUNTRY_TOP_10 |> filter(category == "TV") |> arrange(desc(cumulative_weeks_in_top_10)) |>  slice(1)
longest_run %>%
  select(show_title, country_name, cumulative_weeks_in_top_10)
```
### Question 6: Netflix provides over 200 weeks of service history for all but one country in our data set. Which country is this and when did Netflix cease operations in that country?

#### Answer: Netflix ceased operations in **Russia** on February 27th, 2022.
```{r}
COUNTRY_TOP_10 |>
  group_by(country_name) |>
  summarise(
    total_weeks = n_distinct(week),
    most_recent_week = max(week, na.rm = TRUE),
    .groups = "drop"
  ) |>
  filter(total_weeks < 200)
```

### Question 7:  What is the total viewership of the TV show Squid Game? Note that there are three seasons total and we are looking for the total number of hours watched across all seasons.

#### Answer: The total viewership of *Squid Game* is 5,048,300,000 hours!

```{r}
squid_game_total <- GLOBAL_TOP_10 %>%
  filter(show_title == "Squid Game") %>%
  summarize(total_hours = sum(weekly_hours_viewed, na.rm = TRUE))
print(squid_game_total)
```

### Question 8: The movie Red Notice has a runtime of 1 hour and 58 minutes. Approximately how many views did it receive in 2021? Note that Netflix does not provide the weekly_views values that far back in the past, but you can compute it yourself using the total view time and the runtime.

#### Answer: *Red Notice* received approximately 201,732,203 views in 2021.
```{r}
library(lubridate)
library(scales)

# runtime in hours
runtime_hours <- 118 / 60  # 118 minutes = 1.9666667 hours

red_notice_hours_2021 <- GLOBAL_TOP_10 %>%
  filter(show_title == "Red Notice", year(week) == 2021) %>%
  summarize(total_hours = sum(weekly_hours_viewed, na.rm = TRUE)) %>%
  pull(total_hours)

approx_views_2021 <- red_notice_hours_2021 / runtime_hours
print(approx_views_2021)
```
### Question 9: How many Films reached Number 1 in the US but did not originally debut there? What is the most recent film to pull this off?

#### Answer: 
```{r}

```

### Question 10: Which TV show/season hit the top 10 in the most countries in its debut week? In how many countries did it chart?

#### Answer:  
```{r}

```

# Press Release 1: Upcoming Season of Stranger Things
wasn't able to get to it :(
