---
title: "Mini-Project #04: Just the Fact(-Check)s, Ma’am!"
author: "Haley Vargas"
date: today
format:
  html:
    code-fold: true
    code-summary: "Show the code"
    toc: true
    toc-depth: 3
    toc-title: "📚 Contents"
    smooth-scroll: true
execute:
  echo: true
  warning: false
  message: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(httr2)
library(rvest)
library(dplyr)
library(tidyr)
library(stringr)
library(lubridate)
library(purrr)
library(ggplot2)
library(infer)
library(tidyverse)
```

# Introduction 
This report analyzes payroll levels from the **Current Employment Statistics (CES)** survey and their subsequent **revisions** over the years 1979–2025.  
We download the raw payroll levels, extract revision tables for each year, merge them, and then perform:

- **6 descriptive statistics about CES over the past 45 years**
- **4 ggplot2 visualizations of CES estimates and accuracy over the past 45 years**
- **2 infer-based statistical tests**
- **Fact Check of BLS Revisions**

# Data Acquisition: Download CES Total Nonfarm Payroll

```{r}
req <- request("https://data.bls.gov/pdq/SurveyOutputServlet") |>
  req_method("POST") |>
  req_body_form(
    request_action   = "get_data",
    reformat         = "true",
    from_results_page = "true",
    from_year        = "1979",
    to_year          = "2025",
    `Go.x`           = "38",
    `Go.y`           = "16",
    initial_request  = "false",
    data_tool        = "surveymost",
    series_id        = "CES0000000001",
    years_option     = "specific_years"
  )

resp <- req_perform(req)
html <- resp |> resp_body_html()

# ---- 2. Extract the data table ----
tbl <- html |> 
  html_element("table#table0") |> 
  html_table()

# ---- 3. Pivot from wide (Jan–Dec columns) to long format ----
df <- tbl |>
  pivot_longer(
    cols = -Year,
    names_to = "Month",
    values_to = "level"
  )

# ---- 4. Clean values and convert to proper types ----
payroll_data <- df |>
  mutate(
    Month = str_to_title(Month),                # "JAN" → "Jan"
    level = as.numeric(str_replace(level, ",", "")),  # convert text to numeric
    date  = ym(paste(Year, Month))              # "1979 Jan" → 1979-01-01
  ) |>
  drop_na(level) |>                             # remove footnote rows
  select(date, level) |>
  arrange(date)

payroll_data
```

# Extract Revision Tables 
```{r}
# ---------------------------------------------
# Helper: Extract revisions table for one year
# ---------------------------------------------

pull_year_revisions <- function(year, html_doc) {
  
  # Build CSS selector for the table tied to that year
  selector <- paste0("table.regular#", year)
  
  tbl <- html_doc |>
    html_element(selector) |>
    html_element("tbody") |>
    html_table(header = FALSE)
  
  # Return empty df if nothing was found
  if (is.null(tbl) || nrow(tbl) < 1) {
    return(tibble(
      date = as.Date(character()),
      original = numeric(),
      final = numeric(),
      revision = numeric()
    ))
  }
  
  # Extract the first 12 months + convert columns
  parsed <- tbl |>
    slice(1:12) |>
    transmute(
      month_raw = gsub("\\.", "", X1),
      month_raw = trimws(month_raw),
      # needed columns: original estimate, final estimate, and revision column
      orig_val  = gsub(",", "", X3),
      final_val = gsub(",", "", X5),
      revis_val = gsub(",", "", X8),
      # construct date
      date_str  = paste(year, month_raw),
      date      = suppressWarnings(ym(date_str)),
      original  = suppressWarnings(as.numeric(orig_val)),
      final     = suppressWarnings(as.numeric(final_val)),
      revision  = suppressWarnings(as.numeric(revis_val))
    ) |>
    filter(!is.na(date)) |>
    select(date, original, final, revision)
  
  parsed
}

# ---------------------------------------------
# Master function: Collect all revision years
# ---------------------------------------------

compile_revisions <- function() {
  
  raw_resp <- request("https://www.bls.gov/web/empsit/cesnaicsrev.htm") |>
    req_user_agent("Mozilla/5.0 (X11; Linux x86_64)") |>
    req_perform()
  
  html_doc <- raw_resp |> resp_body_html()
  
  yr_range <- 1979:2025
  
  yr_range |>
    map(~ pull_year_revisions(.x, html_doc)) |>
    list_rbind()
}

# Retrieve full dataset
revision_results <- compile_revisions()
```

# Merge Payroll & Revision Data
```{r}
payroll_revisions <- payroll_data |>
  left_join(revision_results, by = "date") |>
  arrange(date)

payroll_revisions
```

#  Data Exploration and Visualization (Exploratory Statistics)
## What and when were the largest revisions (positive and negative) in CES history?
```{r}
largest_revisions <- payroll_revisions |>
  filter(!is.na(revision)) |>
  slice_max(revision, n = 1, with_ties = FALSE) |>
  mutate(type = "Largest Positive Revision") |>
  bind_rows(
    payroll_revisions |>
      filter(!is.na(revision)) |>
      slice_min(revision, n = 1, with_ties = FALSE) |>
      mutate(type = "Largest Negative Revision")
  )

largest_revisions
```

The largest negative revision of employment was seen in March 1st of 2020, with -672 thousand jobs while the largest positive revision was seen in November 1st of 2021, with 437 thousand jobs.

## Fraction of revisions that were positive by year
```{r}
positive_fraction_year <- payroll_revisions |>
  mutate(pos = revision > 0) |>
  filter(!is.na(revision)) |>
  group_by(year = year(date)) |>
  summarize(
    percent_positive = mean(pos) * 100,
    n_months = n()
  )

positive_fraction_year
```


## Average revision size (absolute and percentage of level)
```{r}
avg_revision_stats <- payroll_revisions |>
  filter(!is.na(revision), !is.na(level)) |>
  summarize(
    avg_revision = mean(revision),
    avg_abs_revision = mean(abs(revision)),
    avg_pct_revision = mean(abs(revision) / level) * 100
  )

avg_revision_stats
```
On average, the revised values changed by 11.2 thousand, with an absolute average shift of 56.7 thousand, representing about a 4.82% revision relative to the original level.

## Relative revision magnitude over time (absolute revision / final estimate)
```{r}
relative_revision_trend <- payroll_revisions |>
  filter(!is.na(final), !is.na(revision)) |>
  mutate(rel_revision = abs(revision) / final) |>
  group_by(year = year(date)) |>
  summarize(
    avg_rel_revision = mean(rel_revision, na.rm = TRUE)
  )

relative_revision_trend
```

## Average revision by month (seasonality patterns)
```{r}

revision_by_month <- payroll_revisions |>
  filter(!is.na(revision)) |>
  mutate(month = month(date, label = TRUE, abbr = TRUE)) |>
  group_by(month) |>
  summarize(
    avg_revision = mean(revision),
    avg_abs_revision = mean(abs(revision))
  )

revision_by_month
```
The average revisions by month shows clear seasonal patterns in revision size. For instance, September shows the largest revision of ≈80.2k jobs, followed by April (≈69.4k), and March (≈64.6k) while February (≈43.7k) and January (≈48.2k) show the smallest. It demonstrates larger spring and early-fall revision adjustments.

## Summary statistics for CES level and revision, 45-year window
```{r}
ces_summary <- payroll_revisions |>
  summarize(
    min_level = min(level, na.rm = TRUE),
    max_level = max(level, na.rm = TRUE),
    mean_level = mean(level, na.rm = TRUE),
    sd_level = sd(level, na.rm = TRUE),
    mean_revision = mean(revision, na.rm = TRUE),
    sd_revision = sd(revision, na.rm = TRUE)
  )

ces_summary
```
There is a positive mean revision of ~11.24 thousand jobs meaning there is employment growth.

#  Visualizations

# Largest Positive & Negative Revisions Over Time
```{r}
ggplot(payroll_revisions, aes(x = date, y = revision)) +
  geom_line() +
  geom_point(data = largest_revisions, aes(x = date, y = revision), 
             color = "red", size = 3) +
  labs(
    title = "CES Revisions Over Time (1979–2025)",
    subtitle = "Red points indicate largest upward and downward revisions",
    x = "Year",
    y = "Revision (Jobs)"
  ) +
  theme_minimal()
```

#  Fraction of Positive Revisions by Year
```{r}
ggplot(positive_fraction_year, aes(x = year, y = percent_positive)) +
  geom_col() +
  labs(
    title = "Fraction of Positive CES Revisions by Year",
    subtitle = "Values closer to 1 = more upward revisions",
    x = "Year",
    y = "Positive Revision Fraction"
  ) +
  theme_minimal()
```

# Relative Revision Magnitude Over Time
```{r}
ggplot(relative_revision_trend, aes(x = year, y = avg_rel_revision)) +
  geom_line() +
  labs(
    title = "Relative CES Revision Magnitude (1979–2025)",
    subtitle = "Absolute revision amount divided by final estimate",
    x = "Year",
    y = "Relative Revision"
  ) +
  theme_minimal()
```

# Average Revision by Month
```{r}
ggplot(revision_by_month, aes(x = month, y = avg_abs_revision)) +
  geom_col() +
  labs(
    title = "Average Absolute CES Revision by Month",
    subtitle = "Seasonality of revision magnitude",
    x = "Month",
    y = "Average Absolute Revision (Jobs)"
  ) +
  theme_minimal()
```

# Statistical Tests

```{r}

payroll_revisions <- payroll_revisions |>
  mutate(
    # logical versions
    neg_revision      = revision < 0,
    large_revision    = abs(revision / level) > 0.01,     # >1% revision
    post_2000         = year(date) >= 2000,
    post_2020         = year(date) >= 2020,
    
    # categorical versions for prop_test()
    neg_cat   = if_else(neg_revision, "negative", "non_negative"),
    large_cat = if_else(large_revision, "large", "small")
  )
```


# Has the fraction of negative revisions increased post-2000?

```{r}
prop_test(
  payroll_revisions,
  response = neg_cat,
  success  = "negative",
  explanatory = post_2000,
  alternative = "greater"
)
```
The p value is greater than 0.05, specifically, 0.186 which shows there is no significant increase of negative revisions post-2000. 

#  Did the average magnitude of revisions increase after 2020?
```{r}
payroll_revisions |>
  mutate(abs_revision = abs(revision)) |>
  t_test(abs_revision ~ post_2020,
         order = c("TRUE", "FALSE"),
         alternative = "greater")

```
**p = 0.01, p < 0.05**
There is statistically significant evidence that the average absolute revision is larger post-2020 than in pre-2020 years. 